---
title: "Handling Platform Differences in Snapshots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Handling Platform Differences in Snapshots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(resultcheck)
```

## The problem: floating-point noise across platforms

Numerical computations in R can produce results that differ at very high decimal
places depending on the platform, compiler, or the version of libraries like
BLAS/LAPACK.  For example, a centering attribute from `scale()` might be
`2.41e-17` on one machine and `1.63e-17` on another — both are effectively zero,
but a naïve snapshot comparison treats them as different.

`resultcheck` provides two complementary mechanisms to handle this:

1. **`[ignored]` line markers** — mark specific lines in a snapshot file as
   known discrepancies that should always be skipped.
2. **`snapshot.precision` in `resultcheck.yml`** — round all floating-point
   numbers to a fixed number of decimal places before comparing or storing
   snapshots.

---

## Feature 1: `[ignored]` line markers

### How it works

Open any `.md` snapshot file under `_resultcheck_snapshots/` and replace the
entire content of a volatile line with the literal text `[ignored]`.  On every
subsequent run, that line position is masked before the comparison takes place,
so differences there will never trigger a failure.  When you update a snapshot
interactively, the `[ignored]` markers are automatically preserved in the new
file.

### Step-by-step example

Suppose your `PCA.R` script produces a snapshot that includes lines like:

```
  ..- attr(*, "scaled:center")= num 2.41e-17
  ..- attr(*, "scaled:center")= num 2.98e-17
```

and these values fluctuate slightly between platforms.  Open the snapshot file
at `_resultcheck_snapshots/PCA/panel_data_pca.md` and change those specific
lines to:

```
[ignored]
[ignored]
```

Save the file.  The snapshot now looks like this (abbreviated):

```markdown
# Snapshot: data.frame

## Structure
'data.frame':  1071 obs. of  30 variables:
 ...
 $ us : num [1:1071, 1] NA -0.423 -0.697 ...
[ignored]
  ..- attr(*, "scaled:scale")= num 1.22
 $ eu : num [1:1071, 1] NA -0.783 -0.711 ...
[ignored]
  ..- attr(*, "scaled:scale")= num 1.37
```

From now on, `run_in_sandbox("code/PCA.R", sandbox)` will not raise an error
regardless of what those two lines contain in any new run, while every other
line continues to be verified.

### Rules

* The **entire line** must be `[ignored]` (leading/trailing whitespace is
  stripped before the check, so indentation does not matter).
* Lines are matched by **position** within the snapshot file, so adding or
  removing other lines shifts the positions.  If the overall structure of the
  output changes materially, update the snapshot interactively first, then
  re-apply the `[ignored]` markers to the new file.
* `[ignored]` markers are **preserved automatically** when you update a
  snapshot: `resultcheck` copies each `[ignored]` line from the stored snapshot
  into the freshly generated one before writing it to disk.

---

## Feature 2: Numeric precision in `resultcheck.yml`

### How it works

Add a `snapshot.precision` key to the `resultcheck.yml` file at the root of
your project.  Before storing or comparing a snapshot, every floating-point
number in the text representation is rounded to that many decimal places.
Integer-like tokens (e.g. index ranges such as `[1:1071]`) are left unchanged.

```yaml
# resultcheck.yml
snapshot:
  precision: 10
```

With `precision: 10`, both `2.41e-17` and `1.63e-17` round to `0`, so the
snapshots compare as equal.

### Step-by-step example

1. Add `precision` to `resultcheck.yml`:

   ```yaml
   snapshot:
     precision: 10
   ```

2. Delete the outdated snapshot (or run interactively and accept the update
   when prompted) so that the stored file is regenerated with rounded values.

3. Re-run the script via `run_in_sandbox()`.  The comparison will now succeed
   whenever the effective difference is smaller than `0.5 × 10^{-precision}`.

### Choosing a precision value

| Situation | Suggested precision |
|-----------|-------------------|
| Machine-epsilon noise (`1e-15` to `1e-17`) | 10 |
| Monte-Carlo / bootstrap variance | 4–6 |
| Monetary amounts (cents) | 2 |

A higher value is more sensitive (closer to exact comparison); a lower value is
more forgiving.

### Backward compatibility

If your project already has snapshot files stored before `precision` was
configured, `resultcheck` also rounds the **stored** snapshot numbers during
comparison.  This means setting `precision` will immediately suppress
floating-point noise in existing snapshots without needing to regenerate them.

---

## Using both features together

The two features are independent and compose naturally:

* Use **`[ignored]`** when a line's value is entirely unreliable across runs
  (e.g. a random seed, a memory address, or a platform-specific constant that
  carries no scientific meaning).
* Use **`precision`** when values are meaningful but may differ in the last few
  decimal places due to floating-point arithmetic.

For the PCA example in the problem statement, either approach resolves the
failure.  Using `precision: 10` is the lower-maintenance option because it
requires no manual edits to snapshot files; `[ignored]` offers surgical control
when only a handful of lines are volatile while the rest must be exact.
